### 2020-02-26 README.Rmd
Fix issue with too few rings, wip
renamed datasets, many bug fixed, working on example

The goal of zonebuilder is to break up large geographic regions such as cities into manageable zones.
Zoning systems are important in many fields, including transport planning, where we hope to use the zones generated by `zonebuilder` to build the evidence base needed to enable uptake of healthy modes of transport like walking, cycling and renewable electricity-powered public transport. See its website at [zonebuilders.github.io/zonebuilder](https://zonebuilders.github.io/zonebuilder/).

@@

```{r install-gh}
# install.packages("devtools")
devtools::install_github("robinlovelace/zonebuilder")
```

@@

Zonebuilder works with `sf` objects and works well alongside the `sf` package and visualisation packages that support spatial data such as `ggplot2`, `leaflet`, `mapdeck`, `mapview` and `tmap`, the last of which we'll use in the following maps.
Attaching the package provides the example dataset `zb_region`, the geographic boundary of a large city (London, UK):

```{r}
library(zonebuilder)
library(tmap)
tmap_border = tm_shape(zb_region) + tm_borders()
tmap_border
```

The core functions in the package break up geographical space into discrete chunks.
The syntax is designed to be user friendly.
The following command, for example, breaks London into quadrats of equal size (and keeping only the space inside the boundary):

```{r}
x = zb_region
q = zb_quadrat(x, ncol = 4) # break into 4
qtm(q) 
```

Another zoning option is 'doughnuts':

```{r}
tmap_arrange(
  qtm(zb_doughnut(x, n_circles = 3)) + tmap_border,
  qtm(zb_doughnut(x, distance = 3)) + tmap_border
)
```

A new zoning system implemented in the function `zb_zone()` segments doughnuts of gradually increasing width to create concentric segmented annuli (CSA):

```{r}
library(zonebuilder)
x = zb_region
z = zb_zone(x)
z
qtm(z)
```

This function gives you control over a number of parameters, as documented in the help pages and an in-progress academic paper.

```{r}
args(zb_zone)
```

The zones are labeled by default in the following way. The rings are labels with capital letters, where A is the centre. The segments are numbers using clock hours, so 12, 3, 6, and 9 for north, east, south, west respectivly.

```{r}
z = zb_zone(x, n_circles = 4)
tm_shape(z) +
  tm_polygons(col = "circle_id", legend.show = FALSE, palette = "magma") + 
  tm_text("label")
```

>>>>

The goal of zonebuilder is to break up large geographic regions such as cities into manageable zones.
Zoning systems are important in many fields, including demographics, economy, health, and transport. The zones have standard configuration, which enabled comparability across cities. See its website at [zonebuilders.github.io/zonebuilder](https://zonebuilders.github.io/zonebuilder/).

@@

```{r install-gh, eval=FALSE}
# install.packages("devtools")
devtools::install_github("robinlovelace/zonebuilder")
```

@@

Zonebuilder works with `sf` objects and works well alongside the `sf` package and visualisation packages that support spatial data such as `ggplot2`, `leaflet`, `mapdeck`, `mapview` and `tmap`, the last of which we'll use in the following maps.
Attaching the package provides the example datasets `london_area` and `london_cent`, the geographic boundary and the centre of London:

```{r}
library(zonebuilder)
library(tmap)
tm_shape(london_area) + tm_borders() + tm_shape(london_cent) + tm_dots("red")
```

The main function `zb_zone` breaks this geographical scale into zones. The default settings follow the **ClockBoard** configuration:

```{r}
library(zonebuilder)
library(tmap)
london_zones = zb_zone(london_area, london_cent)
zb_plot(london_zones)
```

The idea behind this zoning system is based on the following principles:

* Most cities have a centre, the 'heart' of the city. Therefore, the zones are distributed around the centre.
* Typically, the population is much denser in and around the centre and also the traffic intensity is higher. Therefore, the zones are smaller in and around the centre.
* The rings (so A, B, C, D, etc) reflect the proximity to the centre point. The distances from the outer borders of the rings A, B, C, D, etc. follow the triangular number sequence 1, 3, 6, 10, etc. This means that in everyday life use, within zone A everything is in walking distance, from ring B to the centre requires a bike, from zone C and further to the centre typically requires public transport. 
* Regarding direction relative to the centre, we use the clock analogy, since most people are familiar with that. So each ring (annuli) is divided into 12 segments, where segment 12 is directed at 12:00, segment 1 at 1:00 etc.

The package `zonebuilder` does not only create zoning systems based on the CloadBoard layout as illustrated below.

The lower-level fuinctionm

The functions `zb_doughnuts` generates rings:

```{r}
tmap_border = tm_shape(london_area) + tm_borders()
tmap_arrange(
  qtm(zb_doughnut(london_area, london_cent, n_circles = 5), title = "Doughnuts") + tmap_border,
  qtm(zb_segment(london_cent, n_segments = 20), title = "Segments") + tmap_border
)
```

```{r}
tmap_arrange(
  qtm(zb_zone(london_area, london_cent, n_segments = 1, n_circles = 5), title = "Doughnuts") + tmap_border,
  qtm(zb_zone(london_area, london_cent, n_segments = 12, n_circles = 1, segment_center = TRUE), title = "Segments") + tmap_border,
  qtm(zb_quadrat(london_area, london_cent, ncol = 4), title = "Quadrants") + tmap_border
)
```
